### 2019-08-23 manuscript/manuscript.tex
ms updates

\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\V}{\text{V}}
\newcommand{\cf}{\emph{cf.} }
\DeclareMathOperator{\var}{Var}
\DeclareMathOperator{\cov}{Cov}
\DeclareMathOperator{\T}{{\mathrm{T}}}
\newcommand{\vect}[1]{\mathbf{#1}}
\newcommand{\nssh}{SSH_n}

@@

Averaging across replicate populations, we find positive temporal covariances
that are statistically significant (p < XXX) consistent with linked selection
acting to affect allele frequency changes over very short time periods. We
visualize these covariances in Figure \label{figure-1} (A), which depicts the
temporal covariances through time, for each of the five rows covariance matrix.
Each row represents the temporal covariance $\cov(\Delta p_s, \Delta p_t)$,
between some initial reference generation $s$ (the row of the matrix), and some
later timepoint $t$ (the column of the matrix). For each row, the covariances
at first are positive, and then decay towards zero as expected when directional
selection affects linked variants' frequency trajectories until ultimately
linkage disequilibrium and additive genetic variance for decay
\parencite{Buffalo2019-io}.

we can calculate the temporal
covariances between an initial starting reference generation ($s = \{0, 10, 20,
\ldots, 50\})$.

@@

With frequency collected at $\tau$ timepoints across $R$ replicate populations
at $L$ loci, we have an $R \times \tau \times L$ array $\mathbf{F}$ of allele
frequencies and $\mathbf{D}$ array of sequencing depths. We calculate the

change the
dimensionality of

>>>>

\renewcommand{\P}{\mathbb{P}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\V}{\text{V}}
\newcommand{\cf}{\emph{cf.} }
\DeclareMathOperator{\var}{Var}
\DeclareMathOperator{\cov}{Cov}
\DeclareMathOperator{\flt}{\mathrm{flat}}
\DeclareMathOperator{\T}{{\mathrm{T}}}
\newcommand{\vect}[1]{\mathbf{#1}}
\newcommand{\nssh}{SSH_n}

@@

Averaging across replicate populations, we find positive temporal covariances
that are statistically significant (p < XXX) consistent with linked selection
acting to affect allele frequency changes over very short time periods. We
visualize these covariances in Figure \label{figure-1} (A), which depicts the
temporal covariances through time, for each of the five rows covariance matrix.
Each row represents the temporal covariance $\cov(\Delta p_s, \Delta p_t)$,
between some initial reference generation $s$ (the row of the matrix), and some
later timepoint $t$ (the column of the matrix). For each row, the covariances
at first are positive, and then decay towards zero as expected when directional
selection affects linked variants' frequency trajectories until ultimately
linkage disequilibrium and additive genetic variance for decay
\parencite{Buffalo2019-io}. Note that per replicate, the signal is a bit
noisier; see Supplementary Figures XXX.

While the presence of positive temporal covariances is consistent with linked
selection affecting allele frequencies over time, this measure not easily
interpretable. Alternatively, we can quantify the impact of linked selection
through the ratio of total covariance in allele frequency change to the total
variance in allele frequency change. Since the total variation in allele
frequency change can be decomposed into variance and covariance components,
$\var(p_t - p_0) = \sum_{i \ne j} \cov(\Delta p_i, \Delta p_j) / \var(p_t -
p_0)$, and the covariances are zero when only drift acts, this is a
conservative measure of how much of the variance in allele frequency change is
caused by linked selection \parencite{Buffalo2019-io}. Since timepoints every
ten generations were sequenced, we cannot directly estimate the temporal
covariances

@@

With frequency collected at $T+1$ timepoints across $R$ replicate populations
at $L$ loci, we have $\mathbf{F}$ of allele frequencies, $\mathbf{D}$ array of
sequencing depths, and a $\mathbf{N}$ array of the number of individuals
sequenced, each of dimension $R \times (T+1) \times L$. We calculate the array
$\mathbf{\Delta F}$ which contains the allele frequency changes between
adjacent generations, and has dimension $R \times T \times L$. The operation
$\flt(\mathbf{\Delta}\mathbf{F})$ flattens this array to a $(R \cdot
T) \times L$ matrix, such that rows are grouped by replicate, e.g. for
timepoint $t$, replicate $r$, and locus $l$ and allele frequencies $p_{t, r}$,
for a single locus the entries are

% \begin{align}
%   \mathbf{\Delta F} &= 
%                     &\begin{bmatrix} 
%     p_{1, 0, 0} - p_{0, 0, 0} & p_{2, 0, 0} - p_{1, 0, 0} & \ldots & p_{1, 1, 0} - p_{0, 1, 0} & p_{2, 1, 0} - p_{1, 1, 0} & \ldots & p_{T+1, R, 0} - p_{T, R, 0}  \\
%     p_{1, 0, 1} - p_{0, 0, 1} & p_{2, 0, 1} - p_{1, 0, 1} & \ldots & p_{1, 1, 1} - p_{0, 1, 1} & p_{2, 1, 1} - p_{1, 1, 1} & \ldots & p_{T+1, R, 1} - p_{T, R, 1}  \\
%     \vdots & \vdots & \ddots & \vdots & \vdots & \ddots & \vdots  \\
%     p_{1, 0, L} - p_{0, 0, L} & p_{2, 0, L} - p_{1, 0, L} & \ldots & p_{1, 1, L} - p_{0, 1, L} & p_{2, 1, L} - p_{1, 1, L} & \ldots & p_{T+1, R, L} - p_{T, R, L}  \\
%   \end{bmatrix} 
% \end{align}

\begin{align}
    \flt(\mathbf{\Delta F}) &=
                    &\begin{bmatrix} 
    \Delta p_{1, 0, 0} & \Delta p_{2, 0, 0} & \ldots & \Delta p_{1, 1, 0} & \Delta p_{2, 1, 0} & \ldots & \Delta p_{T, R, 0}  \\
    \Delta p_{1, 0, 1} & \Delta p_{2, 0, 1} & \ldots & \Delta p_{1, 1, 1} & \Delta p_{2, 1, 1} & \ldots & \Delta p_{T, R, 1}  \\
    \vdots & \vdots & \ddots & \vdots & \vdots & \ddots & \vdots  \\
    \Delta p_{1, 0, L} & \Delta p_{2, 0, L} & \ldots & \Delta p_{1, 1, L} & \Delta p_{2, 1, L} & \ldots & \Delta p_{T, R, L}  \\
  \end{bmatrix} 
\end{align}

where each $\Delta p_{t, r, l} = p_{t+1, r, l} - p_{t, r, l}$. Then, the sample
temporal-replicate covariance matrix $\mathbf{Q}'$ calculated on
$\flt(\mathbf{\Delta F})$ is a $(R \cdot T) \times (R \cdot T)$
matrix. Given the bias of the sample covariance of allele frequency changes, we
calculated an expected bias matrix $\mathbf{B}$, averaging over loci,

\begin{align}
  \mathbf{B} = \frac{1}{L} \sum_{l=1}^L \frac{\mathbf{h}_l}{2} \circ \left( \frac{1}{\mathbf{d}_l} + \frac{1}{2\mathbf{n}_l} + \frac{1}{2\mathbf{d}_l \circ \mathbf{n}_l} \right)
\end{align}

where $\circ$ denotes elementwise product, and $\mathbf{h}_l$, $\mathbf{d}_l$,
and $\mathbf{n}_l$, are rows corresponding to locus $l$ of the unbiased
heterozygosity arrays $\mathbf{H}$, depth matrix $\mathbf{D}$, and number of
diploids matrix $\mathbf{N}$. The unbiased $R \times (T+1) \times L$
heterozygosity array can be calculated as

\begin{align}
  \mathbf{H} = \frac{2 \mathbf{D} \circ \mathbf{N} }{ (\mathbf{D}-1) \circ (\mathbf{N} -1)} \circ \mathbf{F} \circ (1-\mathbf{F})
\end{align}

Thus, $\mathbf{B}$ is a $R \times (T+1)$ matrix. From XXX, we know that the
temporal variances and covariances require a correction,

\begin{align}
  \mathbf{Q} = \mathbf{Q}' - \mathrm{diag}(\flt(\mathbf{B}_{\cdot,2:(T+1)})) - \mathrm{diag}(\flt(\mathbf{B}_{\cdot, 1:T})) + \mathrm{offdiag}_1(\flt(\mathbf{B}_{\cdot, 2:T})) + \mathrm{offdiag}_{-1}(\flt(\mathbf{B}_{\cdot, 2:T}))
\end{align}

where $\mathrm{diag}(\mathbf{x})$ is an operation that takes the vector
$\mathbf{x}$ and places them along the diagonal of a matrix. Similarly,
$\mathrm{offdiag}_{k}(\mathbf{x})$ places vector $\mathbf{x}$ along the
offdiagonal where for row $i$, and column $j$, $j-i = k$. We represent subset
of columns $s$ through $t$ of a matrix $\mathbf{B}$ as $\mathbf{B}_{\cdot,
s:t}$.
